---
title: "Practical Machine Learning"
output: html_document
---

This document describes the process of training, testing and validating a prediction model for predicting activity classes based on sensor data.

## Getting the data

Downloading the sensor data:

```{r}
#download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", "pml-training.csv", method="curl")
#download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", "pml-testing.csv", method="curl")
```

## Loading the data

Load the data set with training data and the set with the cases we want to predict. We consider empty strings, "NA" and "#DIV/0!" as missing values.

```{r}
projectData <- read.csv("pml-training.csv", na.strings = c("NA", "", "#DIV/0!"))
projectQuestions <- read.csv("pml-testing.csv", na.strings = c("NA", "", "#DIV/0!"))
```

## Cleaning data

We remove the useless variables: NA's and user_name, raw_timestamp_part_1.
```{r}
empty <- subset(names(projectQuestions), 
                apply(projectQuestions, 2, function(x){ all(is.na(x) | x == 0)}))
useless <- c(empty, "X", "user_name", "raw_timestamp_part_1", "raw_timestamp_part_2", "cvtd_timestamp", "new_window", "num_window")
```

And filter them out of the data sets:

```{r}
projectData <- projectData[, !names(projectData) %in% useless]
projectQuestions <- projectQuestions[, !names(projectQuestions) %in% useless]
```

## Available data

The data consists of the following type of useful sensor data:

* {gyros,accel,magnet}\_{belt,arm,dumbbell,forearm}\_{x,y,z}
* {roll,yaw,pitch,total_accel}\_{belt,arm,dumbbell,forearm}

## Testing and training set
Because the testing data had no classe variable, we split the pml-training set in a test set (25%) and a training set (75%).
```{r partition, message=FALSE, results='hide'}
library(caret)
set.seed(1234)
inTrain <- createDataPartition(projectData$classe, p=0.75, list=F)
training <- projectData[inTrain,]
testing <- projectData[-inTrain,]
```

## Creating a prediction model

Create a model using a random forest with 300 trees

```{r rf, message=FALSE}
library(randomForest)
modelFit <- randomForest(classe ~ ., data=training, ntree=300)
```

Predict the outcome of the test set using the model.
```{r}
pred <- predict(modelFit, newdata=testing)
```

Validate the outcome op the prediction on the test set.

```{r}
confusionMatrix(pred, testing$classe)
```

The accuracy of the model is 99,6% which is enough to use it for predicting the outcome of the project questions.

```{r}
finalPred <- predict(modelFit, newdata=projectQuestions)
```
